Low load operating protocol investigation of a 620MWe power boiler using an indirectly coupled process model

### 2021-09-25 PAPER_2/PAPER_2_REV1.tex
Update 25/09/2021

The model was simulated for a boiler load of 32 $\%$  with 6 various firing combinations. To establish the accuracy of the CFD modelling approach a validation study was conducted for 100 $\%$, 80 $\%$ and 60 $\%$ boiler loads and compared to plant measurements to quantify the model accuracy, this is highlighted in section \ref{sec_model_valid} prior to the results of the low-load study.

@@

The CFD model was further validated by comparing the $CO_{ppm}$ and $X_{O_{2}}$ measurements against the CFD results. The probe measurements were taken at a furnace height of 37.5 [$m$] near the center of the boiler during a full load (100\% MCR) operating conditions. The probe is inserted from the side walls to a depth of 4.5 [$m$], measurements were taken every 0.5 [$m$].\\
\begin{figure}[h!]
\centering
\subfigure[]{\includegraphics[width = 0.45\textwidth, height =4cm ]{COPPM_VALID}}
\hspace{5mm}
\subfigure[]{\includegraphics[width = 0.45\textwidth, height =4cm ]{XO2_VALID}}
\caption{Experimentally calculated $CO$ and $O_2$ concentration predictions}
\label{fig_probe_valid}
\end{figure}

@@

Explain the investigation
table with case descriptions
Need flownex model and process modelling description
\begin{figure}
\centering
\includegraphics[scale = 0.45]{TEMP_KEY} [$K$]\\
\subfigure[]{\includegraphics[width=0.32\textwidth]{BOT_TEMP}}
\subfigure[]{\includegraphics[width=0.32\textwidth]{MID_TEMP}}
\subfigure[]{\includegraphics[width=0.32\textwidth]{FBRM_TEMP}}\\
\subfigure[]{\includegraphics[width=0.32\textwidth]{BOT05_TEMP}}
\subfigure[]{\includegraphics[width=0.32\textwidth]{MID05_TEMP}}
\subfigure[]{\includegraphics[width=0.32\textwidth]{FBRM05_TEMP}}
\caption{Temperature fields for cases 1 through 6 [(a)-(f)]}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[scale = 0.45]{VEL_KEY} [$m/s$]\\
\subfigure[]{\includegraphics[width=0.32\textwidth]{BOT_VEL}}
\subfigure[]{\includegraphics[width=0.32\textwidth]{MID_VEL}}
\subfigure[]{\includegraphics[width=0.32\textwidth]{FBRM_VEL}}\\
\subfigure[]{\includegraphics[width=0.32\textwidth]{BOT05_VEL}}
\subfigure[]{\includegraphics[width=0.32\textwidth]{MID05_VEL}}
\subfigure[]{\includegraphics[width=0.32\textwidth]{FBRM05_VEL}}
\caption{Velocity fields for cases 1 through 6 [(a)-(f)]}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[scale = 0.45]{HEATFLUX_KEY} [$W/m^2$]\\
\subfigure[]{\includegraphics[scale = 0.29]{BOT_HEATFLUX}}
\hspace{5mm}
\subfigure[]{\includegraphics[scale = 0.35]{MID_HEATFLUX}}
\hspace{5mm}
\subfigure[]{\includegraphics[scale = 0.35]{FBRM_HEATFLUX}}\\
\subfigure[]{\includegraphics[scale = 0.35]{BOT05_HEATFLUX}}
\hspace{5mm}
\subfigure[]{\includegraphics[scale = 0.35]{MID05_HEATFLUX}}
\hspace{5mm}
\subfigure[]{\includegraphics[scale = 0.35]{FBRM05_HEATFLUX}}
\caption{Heat fluxes profiles for cases 1 through 6 [(a)-(f)]}
\end{figure}

\begin{figure}
\centering
\includegraphics[scale=0.45]{CO_MF_KEY} [$Y_{CO}$] \\
\subfigure[]{	\includegraphics[width=0.18\textwidth, height = 7cm]{BOT_ISO_COPPM_F}
				\includegraphics[height = 7cm]{BOT_ISO_COPPM_S}}
\subfigure[]{	\includegraphics[width=0.18\textwidth, height = 7cm]{MID_ISO_COPPM_F}
				\includegraphics[height = 7cm]{MID_ISO_COPPM_S}}
\subfigure[]{	\includegraphics[width=0.18\textwidth, height = 7cm]{FBRM_ISO_COPPM_F}
				\includegraphics[height = 7cm]{FBRM_ISO_COPPM_S}}\\
\subfigure[]{	\includegraphics[width=0.18\textwidth, height = 7cm]{BOT05_ISO_COPPM_F}
				\includegraphics[height = 7cm]{BOT05_ISO_COPPM_S}}
\subfigure[]{	\includegraphics[width=0.18\textwidth, height = 7cm]{MID05_ISO_COPPM_F}
				\includegraphics[height = 7cm]{MID05_ISO_COPPM_S}}
\subfigure[]{	\includegraphics[width=0.18\textwidth, height = 7cm]{FBRM05_ISO_COPPM_F}
				\includegraphics[height = 7cm]{FBRM05_ISO_COPPM_S}}\\
\caption{CO mass fraction ($Y_{CO}$) concentrations for cases 1 through 6 [(a)-(f)] on a temperature iso-surface of 1600 K}
\end{figure}

\begin{figure}[h!]
\centering
\subfigure[]{\includegraphics[scale = 0.36]{SYMMETRY_PROBE_COPPM}}
\hspace{5mm}
\subfigure[]{\includegraphics[scale = 0.36]{OFFSET_PROBE_COPPM}}\\
\subfigure[]{\includegraphics[scale = 0.35]{SYMMETRY_PROBE_XO2}}
\hspace{5mm}
\subfigure[]{\includegraphics[scale = 0.35]{OFFSET_PROBE_XO2}}

@@

\begin{table}[h!]
\centering
\caption{Process model control parameters}
\label{fuel}
{\tabulinesep=1.2mm
\begin{tabularx}{\linewidth}{p{0.45\linewidth} XXXXXX}
\hline
&\multicolumn{6}{c}{Cases}\\
 & \textbf{1} & \textbf{2} & \textbf{3}& \textbf{4}&\textbf{5}&\textbf{6}\\
\hline
\textbf{Main steam flow rate} 	[$kg/s$]		&166&163&170	&171&169&163\\
\textbf{Main steam exit temp} 	[$^{\circ}C$]	&535&535&535	&535&535&535\\
\textbf{RH steam flow rate} 	[$kg/s$]		&150&148&154	&154&153&147\\
\textbf{RH steam exit temp} 	[$^{\circ}C$]	&535&535&535	&535&520&490\\
\textbf{Boiler efficiency} 		[$\%$]			&86.2&84.9&88.8	&88.2&86.6&82.2\\
\textbf{Attemperator 1} 		[$kg/s$]			&10.5&15&16		&11.5&9.5&12\\
\textbf{Attemperator 2} 		[$kg/s$]			&5&8&5			&4&5&5.5\\
\textbf{Attemperator RH} 		[$kg/s$]			&0.5&1&1		&0&0&0\\
\hline
\end{tabularx}}
\end{table}

\begin{table}[h!]
\centering
\caption{Furnace exit conditions and SH wall temperatures}
\label{fuel}
{\tabulinesep=1.2mm
\begin{tabularx}{\linewidth}{p{0.4\linewidth} XXXXXX}
\hline
&\multicolumn{6}{c}{Cases}\\
 & \textbf{1} & \textbf{2} & \textbf{3}& \textbf{4}&\textbf{5}&\textbf{6}\\
\hline
\multicolumn{7}{l}{\textit{Furnace exit}}\\
\textbf{Exit temperature} [$K$] & 1168 & 1230 & 1215 & 1208 & 1306 & 1298\\
\textbf{Unburnt carbon} [$(\times 10^{-3})\,\%$] & 1.83 & 1.94 & 1.54 & 1.81 & 1.89 & 1.62\\
\multicolumn{7}{l}{\textit{Platen SH}}\\
\textbf{Max wall temperature} [$K$]  &754 & 750 & 765 & 766 & 753 & 763\\
\textbf{Mean wall temperature} [$K$] &726 & 712 & 719 & 727 & 715 & 724\\
\multicolumn{7}{l}{\textit{Pendant SH}}\\
\textbf{Max wall temperature} [$K$]  & 805 & 807 & 811 & 808 & 800 & 802\\
\textbf{Mean wall temperature} [$K$] & 793 & 788 & 792 & 795 & 788 & 784\\
\hline
\end{tabularx}}
\end{table}

\begin{equation}
T_{metal} = T_{wall} - \left(\frac{\dot{q}_{SH}t_{ASH}}{\lambda_{ASH}}\right)
\end{equation}

@@

\clearpage
\section{Conclusions}

\section*{Acknowledgements}
The authors would like to thank the Eskom EPPEI program for financially supporting the present study and acknowledge the computational resources provided by the Centre for High Performance Computing (CHPC), South Africa.

>>>>

The model was simulated for a boiler load of 32 $\%$  with 6 various firing combinations. To establish the accuracy of the CFD modelling approach a validation study was conducted for 100 $\%$, 80 $\%$ and 60 $\%$ maximum continuous rating (MCR) loads and compared to actual plant measurements to quantify the model accuracy, this is highlighted in section \ref{sec_model_valid} prior to the results of the low-load study.

@@

The CFD model was further validated by comparing the $CO_{ppm}$ and $X_{O_{2}}$ measurements against the CFD results. The probe measurements were taken at a furnace height of 37.5 [$m$] near the center of the boiler during a full load (100\% MCR) operating conditions. The probe is inserted from the side walls to a depth of 4.5 [$m$], measurements were taken every 0.5 [$m$].\\
\begin{figure}[h!]
\centering
\subfigure[]{\includegraphics[width = 0.45\textwidth, height =4cm ]{COPPM_VALID}}
\hspace{5mm}
\subfigure[]{\includegraphics[width = 0.45\textwidth, height =4cm ]{XO2_VALID}}
\caption{Experimentally calculated $CO_{ppm}$ (a) and $X_{O_{2}}$ (b) concentration predictions}
\label{fig_probe_valid}
\end{figure}

@@

Figures \ref{fig_cfd_temp} and \ref{fig_cfd_velo} illustrate the temperature and velocity profiles for the various firing arrangements. With only the middle burner rows firing (Cases 2 and 5) there a is substantial cold region located in the lower half the furnace, resulting in the lowest heat uptake of the furnace for all the cases. This is further exaggerated when considering the Table \ref{tbl_process_parameters} with the mid-firing arrangement producing the lowest steam flow rates resulting in the lowest boiler efficiencies. The bottom firing arrangement (Cases 1 and 4) results in fireball located in the bottom half on the burner, while the mixed firing arrangement (Cases 3 and 6) results is more even distribution of the fireball in the furnace domain. This leads to the highest steam generation rate and boiler efficiency as shown in Table \ref{tbl_process_parameters}.

Figure \ref{fig_cfd_heat_flux} illustrates the heat flux profiles for the simulated cases. Both cases 1 and 4 highlight high heat flux zones near the burner inlets which, in the presence of high temperatures and incomplete combustion near these regions, can lead to high-temperature corrosion \cite{Du2017}. Cases 2 and 5 show that most of the unit heat fluxes are absorbed in the upper half of the boiler, contrary to cases 1 and 4, where the lower half has a higher heat flux absorption, and cases 3 and 6 showing an even distribution throughout the furnace.

The works of Dugum et al \citep{Du2017}, highlights the issue of high-temperature corrosion caused by significant levels of $CO$ ($X_CO$ 0.01 - 0.1) and no-free $O_2$ near regions of high wall temperatures. For low-load operation this phenomena becomes important to avoid since combustion instability can lead to these ideal circumstances. Figure \ref{fig_cfd_coppm} shows the $CO$ molar concentration in the domain located on an iso-surface set 1600 $[K]$. Cases 1 and 4 illustrate the highest likelihood of high-temperature corrosion occurring near the top half of the furnace hopper, furthermore considering Figures \ref{fig_cfd_coppm}(c) and (d) a lower concentration of oxygen is reported.

Investigating the combustion stability for all the cases, symmetry and offset vertical probe plots (as highlighted in Figure \ref{fig_geometry}) are given in Figure \ref{fig_cfd_probe}. The general trends of the $CO_{ppm}$ correspond with each cases burner firing arrangement, as seen with cases 3 and 6, a mixed firing arrangement, where two peaks are observed in the vicinity of the bottom and middle burner rows. Cases 2 and 5 illustrates the highest $X_{O_{2}}$ concentration in the lower half of the burner since there is minimal combustion occurring. The unburnt carbon content and exit flue-gas temperatures for each case are reported in Table \ref{tbl_cfd_results}. With the use of the middle burners (Cases 2 and 5) the highest exit temperature and unburnt carbon content is observed, since the fireball is located in the upper half of the furnace leading to the least possibility of complete combustion due to the shorter residence time. Cases 3 and 6 exhibit the best characteristics with the least amount of unburnt carbon content being observed. It is important to note the effects of a lower SA mass flow rate has on the furnace exit conditions for non-firing burners which generally leads to a hotter exit temperature as is the case with cases 4 to 6.

The external tube metal temperatures, of Figure \ref{tbl_cfd_results}, were calculated using equation \ref{eqn_t_wall}, which takes into account the temperature drop due to the ash deposit present on platen and final SH.
\begin{equation}\label{eqn_t_wall}
T_{metal} = T_{wall} - \left(\frac{\dot{q}_{SH}t_{ASH}}{\lambda_{ASH}}\right)
\end{equation}
For the platen and final SH the maximum surface temperature are observed for cases 2 and 5. For comparison a 100 \% MCR load the maximum temperatures reported for the platen and final SH are of 500 $[^\circ C]$ and 623 $[^\circ C]$, with the final SH operating in the materials creep range \cite{Laubscher2019b}. Thus,

Using the process model of Figure \ref{fig_flownex} and the results of the CFD simulations the important process control parameters were determined.

\begin{figure}\label{fig_cfd_temp}
\centering
\includegraphics[scale = 0.45]{TEMP_KEY} [$K$]\\
\subfigure[Case 1]{\includegraphics[width=0.32\textwidth]{BOT_TEMP}}
\subfigure[Case 2]{\includegraphics[width=0.32\textwidth]{MID_TEMP}}
\subfigure[Case 3]{\includegraphics[width=0.32\textwidth]{FBRM_TEMP}}\\
\subfigure[Case 4]{\includegraphics[width=0.32\textwidth]{BOT05_TEMP}}
\subfigure[Case 5]{\includegraphics[width=0.32\textwidth]{MID05_TEMP}}
\subfigure[Case 6]{\includegraphics[width=0.32\textwidth]{FBRM05_TEMP}}
\caption{Temperature fields for cases 1 through 6 [(a)-(f)]}
\end{figure}

\begin{figure}[h!]\label{fig_cfd_velo}
\centering
\includegraphics[scale = 0.45]{VEL_KEY} [$m/s$]\\
\subfigure[Case 1]{\includegraphics[width=0.32\textwidth]{BOT_VEL}}
\subfigure[Case 2]{\includegraphics[width=0.32\textwidth]{MID_VEL}}
\subfigure[Case 3]{\includegraphics[width=0.32\textwidth]{FBRM_VEL}}\\
\subfigure[Case 4]{\includegraphics[width=0.32\textwidth]{BOT05_VEL}}
\subfigure[Case 5]{\includegraphics[width=0.32\textwidth]{MID05_VEL}}
\subfigure[Case 6]{\includegraphics[width=0.32\textwidth]{FBRM05_VEL}}
\caption{Velocity fields for cases 1 through 6 [(a)-(f)]}
\end{figure}

\begin{figure}[h!]\label{fig_cfd_heat_flux}
\centering
\includegraphics[scale = 0.45]{HEATFLUX_KEY} [$W/m^2$]\\
\subfigure[Case 1]{\includegraphics[scale = 0.29]{BOT_HEATFLUX}}
\hspace{5mm}
\subfigure[Case 2]{\includegraphics[scale = 0.35]{MID_HEATFLUX}}
\hspace{5mm}
\subfigure[Case 3]{\includegraphics[scale = 0.35]{FBRM_HEATFLUX}}\\
\subfigure[Case 4]{\includegraphics[scale = 0.35]{BOT05_HEATFLUX}}
\hspace{5mm}
\subfigure[Case 5]{\includegraphics[scale = 0.35]{MID05_HEATFLUX}}
\hspace{5mm}
\subfigure[Case 6]{\includegraphics[scale = 0.35]{FBRM05_HEATFLUX}}
\caption{Heat fluxes profiles for cases 1 through 6 [(a)-(f)]}
\end{figure}

\begin{figure}\label{fig_cfd_coppm}
\centering
\includegraphics[scale=0.45]{CO_MF_KEY} [$X_{CO}$] \\
\subfigure[Case 1]{	\includegraphics[width=0.18\textwidth, height = 7cm]{BOT_ISO_COPPM_F}
				\includegraphics[height = 7cm]{BOT_ISO_COPPM_S}}
\subfigure[Case 2]{	\includegraphics[width=0.18\textwidth, height = 7cm]{MID_ISO_COPPM_F}
				\includegraphics[height = 7cm]{MID_ISO_COPPM_S}}
\subfigure[Case 3]{	\includegraphics[width=0.18\textwidth, height = 7cm]{FBRM_ISO_COPPM_F}
				\includegraphics[height = 7cm]{FBRM_ISO_COPPM_S}}\\
\subfigure[Case 4]{	\includegraphics[width=0.18\textwidth, height = 7cm]{BOT05_ISO_COPPM_F}
				\includegraphics[height = 7cm]{BOT05_ISO_COPPM_S}}
\subfigure[Case 5]{	\includegraphics[width=0.18\textwidth, height = 7cm]{MID05_ISO_COPPM_F}
				\includegraphics[height = 7cm]{MID05_ISO_COPPM_S}}
\subfigure[Case 6]{	\includegraphics[width=0.18\textwidth, height = 7cm]{FBRM05_ISO_COPPM_F}
				\includegraphics[height = 7cm]{FBRM05_ISO_COPPM_S}}\\
\caption{CO molar fraction ($X_{CO}$) concentrations for cases 1 through 6 [(a)-(f)] on a temperature iso-surface of 1600 K}
\end{figure}

\begin{figure}[h!]\label{fig_cfd_probe}
\centering
\subfigure[Symmetry vertical plot]{\includegraphics[scale = 0.36]{SYMMETRY_PROBE_COPPM}}
\hspace{5mm}
\subfigure[Offset vertical plot]{\includegraphics[scale = 0.36]{OFFSET_PROBE_COPPM}}\\
\subfigure[Symmetry vertical plot]{\includegraphics[scale = 0.35]{SYMMETRY_PROBE_XO2}}
\hspace{5mm}
\subfigure[Offset vertical plot]{\includegraphics[scale = 0.35]{OFFSET_PROBE_XO2}}

@@

\begin{table}[h!]\label{tbl_process_parameters}
\centering
\caption{Process model control parameters}
\label{fuel}
{\tabulinesep=1.2mm
\begin{tabularx}{\linewidth}{p{0.45\linewidth} XXXXXX}
\hline
&\multicolumn{6}{c}{Cases}\\
 & \textbf{1} & \textbf{2} & \textbf{3}& \textbf{4}&\textbf{5}&\textbf{6}\\
\hline
\textbf{Main steam flow rate} 	[$kg/s$]		&175.8&172.9&180.5&180.2&179.1&184.1 \\
\textbf{Main steam exit temp} 	[$^{\circ}C$]	&535& 535 &535&535 &535& 535\\
\textbf{RH steam flow rate} 	[$kg/s$]		&158.2&155.6&162.5&162.2&161.2&165.6\\
\textbf{RH steam exit temp} 	[$^{\circ}C$]	&524&527&531&520&510&512\\
\textbf{Boiler efficiency} 		[$\%$]			&85.3&84.1&87.9	&87.2&85.9&89.1\\
\textbf{ATT1} 		[$kg/s$]					&10.4&16.9&13.9&7.9&5.5&10.9\\
\textbf{ATT2} 		[$kg/s$]					&1.7&3.8&4.2&3.8&3.67&4.2\\
\textbf{ATT-RH} 		[$kg/s$]				&0.0&0.0&0.0&0.0&0.0&0.0\\
\hline
\end{tabularx}}
\end{table}

\begin{table}[h!]\label{tbl_cfd_results}
\centering
\caption{Furnace exit conditions and SH wall temperatures}
\label{fuel}
{\tabulinesep=1.2mm
\begin{tabularx}{\linewidth}{p{0.4\linewidth} XXXXXX}
\hline
&\multicolumn{6}{c}{Cases}\\
 & \textbf{1} & \textbf{2} & \textbf{3}& \textbf{4}&\textbf{5}&\textbf{6}\\
\hline
\multicolumn{7}{l}{\textit{Furnace exit}}\\
\textbf{Exit temperature} [$K$] & 1168 & 1230 & 1215 & 1208 & 1306 & 1298\\
\textbf{Unburnt carbon} [$(\times 10^{-3})\,\%$] & 1.83 & 1.94 & 1.54 & 1.81 & 1.89 & 1.62\\
\multicolumn{7}{l}{\textit{Platen SH}}\\
\textbf{Max wall temperature} [$^{\circ}C$]  &477 & 492 & 481  & 480 & 493 & 490\\
\textbf{Mean wall temperature} [$^{\circ}C$] &439 & 453 & 446 & 454 & 442 & 451\\
\multicolumn{7}{l}{\textit{Final SH}}\\
\textbf{Max wall temperature} [$^{\circ}C$]  & 602 & 624 & 608 & 595 & 626 & 612\\
\textbf{Mean wall temperature} [$^{\circ}C$] & 520 & 523 & 517 & 512 & 520 & 511\\
\hline
\end{tabularx}}
\end{table}

@@

\clearpage
\section{Conclusions}
ghhhb
jyjjf
\section*{Acknowledgements}
The authors would like to thank the Eskom EPPEI program for financially supporting the present study and acknowledge the computational resources provided by the Centre for High Performance Computing (CHPC), South Africa.
